---
title: "TFCB2020 Capstone Project"
author: "BrendanMcShane"
date: "12/16/2020"
output: html_document
---

Question 3:Which continuous variables are the most correlated, and to what extent? Do these correlations change after subsetting by disease state?

```{r, echo = FALSE, warning = FALSE, message=FALSE}
#Load in all the libraries necessary for this code
library(tidyverse)
#I found the library below containing convenient functions to both calculate and plot correlations
library(corrplot)
#After discussion with Maggie, I also used the following library for its kable function for prettier looking tables
library(knitr)
```

```{r}
df <- read.csv(file = 'data/ds_whole.csv')
#Read in the data
df_cont <- df[, c(3,5,6,8,9)]
#Subset to the continuous variables
df_cont_corr <- cor(df_cont)
#Using the cor() function from corrplot to generate correlation coefficients for each pair of variables
knitr::kable(df_cont_corr)
#make a well-formatted table of the calculated correlation coefficients
```

```{r}
corrplot(df_cont_corr, 
         method = "square", 
         type = "lower", 
         diag = FALSE, 
         tl.pos = "ld", 
         tl.col = "black",
         tl.cex = 1.1)

```

The correlations between these continuous variables show that RNFL4.mean and mean_deviation are the most well correlated, with a Pearson correlation coefficient of ~0.679.
The visualization above displays the correlation coefficients using both a red-blue color scale and by the size of the square within the box. I trimmed off the mirroring of the values, self-pairings,  andchanged the font color to black for better readability.


I'll now repeat these analyses on the data after subsetting by disease state.

First I'll do the non-glaucoma observations:
```{r}
df_healthy_cont <- df[which(df$glaucoma == 0), c(3,5,6,8,9)]
#Here I use the which function to take data only from rows where the value for the variable "glaucoma" is 0, indicating healthy samples
#Afterwards, visualization is the same as above.
df_healthy_cont_corr <- cor(df_healthy_cont)
kable(df_healthy_cont_corr)


```
```{r}
corrplot(df_healthy_cont_corr, 
         method = "square", 
         type = "lower", 
         diag = FALSE, 
         tl.pos = "ld", 
         tl.col = "black",
         tl.cex = 1.1)
```


In the healthy samples, mean deviation and pattern standard deviation are now the strongest correlation, being an anti-correlation with a value of -0.57. It is somewhat interesting that the dominant correlation changes

```{r}
df_disease_cont <- df[which(df$glaucoma == 1), c(3,5,6,8,9)]
#Same type of subsetting by row, except using a value of one, indicating glaucoma diagnosis.
df_disease_cont_corr <- cor(df_disease_cont)
kable(df_disease_cont_corr)

```

```{r}
corrplot(df_disease_cont_corr, 
         method = "square", 
         type = "lower", 
         diag = FALSE, 
         tl.pos = "ld", 
         tl.col = "black")
```


In the disease state, mean_deviation and RNFL mean maintain the highest correlation coefficient, although the strength of the correlation weakens slightly having a value of only 0.56.

#Repoducibility
Hopefully between the comments in the code and the README, my analyses are easy to follow through and reproduce. I'll say that the original authors do a reasonable job describing the dataset and stating which models they use to compare to; however, I could not find the actual code they used for their model. Whether this is my own failing to understand/interpret their work or them not actually making it clear is not apparent to me.





